DisReserv (v1.0): Analytical displacement solution due to reservoir compaction with arbitrary geometry and under arbitrary pressure changes

### 2020-09-18 manuscript/manuscript-reservoir.tex
subsection 3.1

The discrete forward modeling to calculate the displacement and stress fields due to a piecewise-constant distribution of the pressure contrast within a reservoir follows the nucleus-of-strain approach. 
We assume that a nucleus of strain represents an infinitesimal reservoir volume 
element. The displacement solution for a single nucleus of strain in a homogeneous elastic semi-infinite medium (equation \ref{eq:displacement-solution}) will be used as an element of the displacement.
Then we calculate the displacement field due to the pressure contrast of a prism by integrating the nucleus of strain over its volume. The same approach is used to 
compute the stress field due to the prism.

@@

By considering the discrete form of equation \ref{eq:displacement-solution}, the displacement field  $ {\bf{u}}_{i} \equiv  {\bf{u}}(x_{i}, y_{i}, z_{i})$ at an arbitrary point $(x_{i}, y_{i}, z_{i})$ due to the $j$th nucleus of strain at the coordinates 
$(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ will be calculated by
\begin{equation}
{\bf{u}}_{i}  = {\bf{\nabla}} \: 
{\phi_{1}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})} \: + \: {\bf{{\nabla}_{2}}} {\phi_{2}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})} \: . 
\label{eq:displacement_ui}
\end{equation}
In equation \ref{eq:displacement_ui}, the functions 
$\phi_{1} \equiv {\phi_{1}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})}$ and  
$\phi_{2} \equiv {\phi_{1}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, -z^{\prime}_{j})}$
are, respectively, given by 
\begin{equation}
\phi_{1} = - \frac{C_m}{4 \pi}  \frac{\Delta p_{j} \: \:dv_j}{ {R_1}_{ij} }
\label{eq:nucleus-phi1}
\end{equation}
and
\begin{equation}
\phi_{2} = - \frac{C_m}{4 \pi}  \frac{\Delta p_{j} \: \:dv_j}{ {R_2}_{ij} }.
\label{eq:nucleus-phi2}
\end{equation}
In equations \ref{eq:nucleus-phi1} and \ref{eq:nucleus-phi2}, $\Delta p_j$  is the pressure contrast of the $j$th nucleus, $dv_j$ is an infinitesimal element of volume of the $j$th nucleus, and $C_m$ is the uniaxial compaction coefficient (see \citeauthor{Geertsma66}, \citeyear{Geertsma66}; \citeauthor{Tempone10}, \citeyear{Tempone10} and \citeauthor{Munoz&Roehl17}, \citeyear{Munoz&Roehl17}) given by
\begin{equation}
C_m = \frac{1}{E} \: \frac{(1 + \nu) (1  - 2\nu)}{(1-\nu)},
\label{eq:Cm}
\end{equation}
where $E$ is the Young’s modulus.
In equation \ref{eq:nucleus-phi1}, $ {R_1}_{ij}$ is the distance from the $i$th coordinate point of the displacement $ (x_{i}, y_{i}, z_{i})$ to the $j$th coordinate of the nucleus of strain $(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$, i.e.:
\begin{equation}
{R_1}_{ij} = {\sqrt{(x_{i}- x^{\prime}_{j})^{2} + (y_{i} - y^{\prime}_{j})^{2} + 
(z_{i} - z^{\prime}_{j})^{2}}}.
\label{eq:R1}
\end{equation}
In equation \ref{eq:nucleus-phi2}, $ {R_2}_{ij}$ is the distance from the $i$th coordinate point of the displacement $ (x_{i}, y_{i}, z_{i})$ to the $j$th coordinate of the image nucleus $(x^{\prime}_{j}, y^{\prime}_{j}, - z^{\prime}_{j})$, i.e.:
\begin{equation}
{R_2}_{ij} = {\sqrt{(x_{i}- x^{\prime}_{j})^{2} + (y_{i} - y^{\prime}_{j})^{2} + 
(z_{i} + z^{\prime}_{j})^{2}}}.
\label{eq:R2}
\end{equation}

Figure \ref{fig:nucleus_strain} shows a schematic representation of the geometry
of the nucleus of strain problem in a semi-infinite medium. 
The $j$th nucleus of strain is located at the coordinates 
$(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$. 
The $j$th image nucleus is located at the coordinates 
$(x^{\prime}_{j}, y^{\prime}_{j}, - z^{\prime}_{j})$.
The distances from the $i$th coordinate point of the displacement $ (x_{i}, y_{i}, z_{i})$ to the $j$th nucleus of strain and to the $j$th image nucleus are, respectively,  
$ {R_1}_{ij}$ (equation \ref{eq:R1})  and $ {R_2}_{ij}$ (equation \ref{eq:R2}).
The horizontal plane $z = 0$ is called free surface.

Following the discrete form of the displacement solution (equation \ref{eq:displacement_ui}), 
the displacement field at the coordinates $x_i$, $y_i$  and $z_i$  due to the $j$th single nucleus at the coordinates $(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ can be written as:
\begin{equation}
{\bf{u}}_{i} (x_i,y_i,z_i) = {\bf{u_{1}}}_{i} (x_i,y_i,z_i) + {\bf{u_{2}}}_{i} (x_i,y_i,z_i), 
\label{eq:nucleus-ui_total}
\end{equation}
where ${\bf{u_{1}}}_{i} (x_i,y_i,z_i) \equiv {\bf{u_{1}}}_{i} $ is the gradient of the function $\phi_1$ (equation \ref{eq:nucleus-phi1}) 
\begin{equation}
{\bf{u_{1}}}_{i}  = {\bf{\nabla}} \: {\phi_{1}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})} \:
= \frac{A  \: (1 + \nu)}{E} \: {\bf{\nabla}} \bigg( {\frac{1}{{R_1}_{ij}}} \bigg) 
\: \Delta p_{j} \: \:dv_j
\label{eq:nucleus-u1}
\end{equation}
and ${\bf{u_{2}}}_{i} (x_i,y_i,z_i) \equiv {\bf{u_{2}}}_{i} $ is obtained by  applying the operator ${\bf{{\nabla}_{2}}}$ (equation \ref{eq:nabla2}) to the image potential $\phi_2$ (equation \ref{eq:nucleus-phi2})
\begin{equation}
{\bf{u_{2}}}_{i} = {\bf{{\nabla}_{2}}} \: {\phi_{2}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, -z^{\prime}_{j})} \:  
= \frac{A  \: (1 + \nu)}{E} \: 
\bigg[ 
(3 - 4\nu){\bf{\nabla}} \bigg(  {\frac{1}{{R_2}_{ij}}} \bigg)     
+ 2 {\bf{\nabla}} \bigg( z \frac{\partial }{\partial z} {\frac{1}{{R_2}_{ij}}} \bigg)  
- 4(1- \nu){\bf{\hat{z}}} {\bf{{\nabla}^{2}}}  \bigg( {\frac{z}{{R_2}_{ij}}} \bigg)
\bigg]   \: \Delta p_{j}  \:dv_j , 
\label{eq:nucleus-u2}
\end{equation}
where  $A$ is a constant given by: 
\begin{equation}
A = - \frac{C_m E}{4 \pi (1 + \nu)}.  
\label{eq:nucleus-A}
\end{equation}
The elements  of the displacement vector ${\bf{u_{1}}}_{i}$ (equation \ref{eq:nucleus-u1}) are
\begin{equation}
{\bf{u_{1}}}_{i}  = 
\begin{bmatrix} 
{u_{1}}_x \\
{u_{1}}_y \\
{u_{1}}_z
\end{bmatrix}
 = \frac{A  \: (1 + \nu)}{E} \: 
\begin{bmatrix} 
\frac{\partial }{\partial x} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial }{\partial y} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial }{\partial z} {\frac{1}{{R_1}_{ij}}} 
\end{bmatrix}
 \: \Delta p_{j} \: \:dv_j \:  ,  
\label{eq:nucleus-u1-vector}
\end{equation}
where ${u_{1}}_x$, ${u_{1}}_y$ and ${u_{1}}_z$ are the $x-$, $y-$  and $z-$ components
of  ${\bf{u_{1}}}_{i} $ that gives the displacement field at the coordinates $x_i$, $y_i$  and $z_i$  due to the $j$th single nucleus in the infinite space.

The elements  of the displacement vector ${\bf{u_{2}}}_{i} $ (equation \ref{eq:nucleus-u2}) are
\begin{equation}
{\bf{u_{2}}}_{i}  = 
\begin{bmatrix} 
{u_{2}}_x \\
{u_{2}}_y \\
{u_{2}}_z
\end{bmatrix}
 = \frac{A  \: (1 + \nu)}{E} \: \left\{   (3  - 4 \nu)
\begin{bmatrix} 
  \: \frac{\partial }{\partial x} {\frac{1}{{R_2}_{ij}}}  \\
  \: \frac{\partial }{\partial y} {\frac{1}{{R_2}_{ij}}}  \\
 - \frac{\partial }{\partial z} {\frac{1}{{R_2}_{ij}}}  
\end{bmatrix}
+ 2 \: z_{i}
\begin{bmatrix} 
\frac{\partial^{2}  }{\partial x \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_2}_{ij}}} 
\end{bmatrix}
\right\}
\: \Delta p_{j} \: \:dv_j,
\label{eq:nucleus-u2-vector}
\end{equation}
where ${u_{2}}_x$, ${u_{2}}_y$ and ${u_{2}}_z$ are the $x-$, $y-$  and $z-$ components
of  ${\bf{u_{2}}}_{i} $ that gives the correction of the displacements considering  a semi-space (image nucleus solution).

By following \cite{Sharma56} and \cite{Tempone10}, the Beltrami’s equations \citep{Beltrami} and the equilibrium equations must be satisfied to obtain the contribution of the stress field in the half space. 
The stress field  at the coordinates $x_i$, $y_i$ and $z_i$  due to the $j$th single nucleus of strain buried in the half space is given by
\begin{equation}
\mbox{\boldmath$\sigma$}_{i}(x_i,y_i,z_i) = \mbox{\boldmath$\sigma_{1}$}(x_i,y_i,z_i) + 
\mbox{\boldmath$\sigma_{2}$}(x_i,y_i,z_i) \: ,
\label{eq:stress_nucleus}
\end{equation}
where $\mbox{\boldmath$\sigma_{1}$}(x_i,y_i,z_i) \equiv \mbox{\boldmath$\sigma_{1}$}_{i}$ represents the stress in an infinite medium, and 
$\mbox{\boldmath$\sigma_{2}$}(x_i,y_i,z_i) \equiv \mbox{\boldmath$\sigma_{2}$}_{i}$ represents a correction of the stress in a half-space due to an image nucleus.
Besides the Beltrami’s equations \citep{Beltrami} and the equilibrium equations that must be satisfied, the following boundary conditions at the free surface ($z_i = 0$) must be satisfied, i.e.:
\begin{equation}
\mbox{\boldmath$\sigma$}_{i}(x_i,y_i,0) = \mbox{\boldmath$\sigma_{1}$}(x_i,y_i,0) + 
\mbox{\boldmath$\sigma_{2}$}(x_i,y_i,0) \: = \: \mbox{ \boldmath$0$},
\label{eq:null_stress}
\end{equation}
where $\mbox{ \boldmath$0$}$ is the null vector.
The elements  of the stress vector $\mbox{\boldmath$\sigma_{1}$}_{i}$ (equation \ref{eq:stress_nucleus}) are
\begin{equation}
\mbox{\boldmath$\sigma_{1}$}_{i}  = 
\begin{bmatrix} 
{\widehat{xz}_{1}} \\
{\widehat{yz}_{1}}  \\
{\widehat{zz}_{1}} 
\end{bmatrix}
 = A  \: 
\begin{bmatrix} 
\frac{\partial^{2} }{\partial x \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_1}_{ij}}} 
\end{bmatrix}
 \: \Delta p_{j} \: \:dv_j \:  ,  
\label{eq:nucleus-stress1-vector}
\end{equation}
where ${\widehat{xz}_{1}}$, ${\widehat{yz}_{1}}$ and ${\widehat{zz}_{1}}$ are the $x-$, $y-$  and $z-$components
of  $\mbox{\boldmath$\sigma_{1}$}_{i}$ that gives the stress in an infinite medium due to the $j$th nucleus of strain.
The elements  of the stress vector $\mbox{\boldmath$\sigma_{2}$}_{i}$ (equation \ref{eq:stress_nucleus}) are
\begin{equation}
\mbox{\boldmath$\sigma_{2}$}_{i}  = 
\begin{bmatrix} 
{\widehat{xz}_{2}} \\
{\widehat{yz}_{2}}  \\
{\widehat{zz}_{2}} 
\end{bmatrix}
= A  \: 
\left\{
\begin{bmatrix} 
\frac{\partial^{2}  }{\partial x \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_2}_{ij}}} \\
- \frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_2}_{ij}}}   
\end{bmatrix}
+ 2 \: z_{i}
\begin{bmatrix} 
\frac{\partial^{3}  }{\partial x \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial y \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial z^{3}} {\frac{1}{{R_2}_{ij}}} 
\end{bmatrix}
\right\}
\: \Delta p_{j} \: \:dv_j,
\label{eq:nucleus-stress2-vector}
\end{equation}
where ${\widehat{xz}_{2}}$, ${\widehat{yz}_{2}}$ and ${\widehat{zz}_{2}}$ are the $x-$, $y-$  and $z-$components
of  $\mbox{\boldmath$\sigma_{2}$}_{i}$ that gives the correction of the stress considering  a semi-space due to the $j$th image nucleus.
By taking the elements of stress vectors $\mbox{\boldmath$\sigma_{1}$}_{i}$ and 
$\mbox{\boldmath$\sigma_{2}$}_{i}$ at any point on the free surface $(x_i, y_i, z_i = 0)$, we obtain
\begin{equation}
{\widehat{xz}_{1}} + {\widehat{xz}_{2}} 
= {\widehat{yz}_{1}} + {\widehat{yz}_{2}}
= {\widehat{zz}_{1}} + {\widehat{zz}_{2}}  
= 0 \: ,
\label{eq:null_stress_i}
\end{equation}
which satisfies the boundary conditions at the free surface 
(equation \ref{eq:null_stress}).

@@

We parameterized the reservoir as a grid of juxtaposed right rectangular prisms.
Each grid prism has homogeneous pressure contrasts; however, pressure variations among different prisms are allowed. 
To calculate the displacement due to the pressure change in the whole reservoir with this discretization model, we use the solution deduced for a single nucleus of strain in a homogeneous elastic semi-infinite medium (subsection \ref{solution-nucleus}) in the following way. 
First, we assume that the coordinates of the $j$th prism center are the coordinates of a nucleus of strain.
Next, the displacement field calculated at the $i$th coordinates  $(x_i$, $y_i, z_i)$  due to the pressure contrast of the $j$th prism is calculated with a integration over its volume.
Then, from equation \ref{eq:nucleus-ui_total}, the displacement field produced by the $jth$ prism can be written as 
\begin{equation}
{\bf{u}}_{i} (x_i,y_i,z_i) =  
\int_{z_1}^{z_2} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{1}}}_{i} \: \:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j 
\: + \:
\int_{z_1}^{z_2} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{2}}}_{i} \:\:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j.
\label{eq:nucleus-ui_jth_prism}
\end{equation}

In equation \ref{eq:nucleus-ui_jth_prism},  ${\bf{u_{1}}}_{i}$ is the displacement field at the coordinates $x_i$, $y_i$  and $z_i$  due to the $j$th single nucleus in the infinite space (equations \ref{eq:nucleus-u1} and  \ref{eq:nucleus-u1-vector}), the limits of the integrals represent the borders of the $j$th prism modeling the reservoir in the following way: $x_1$ and $x_2$ are their south and north borders; $y_1$ and  $y_2$ are their  west and east borders;  and $z_1$ and $z_2$ are their depths to the top and bottom.
Additionally, ${\bf{u_{2}}}_{i}$ is the displacement field at the coordinates $x_i$, $y_i$  and $z_i$  due to the effect of an image nucleus (equations \ref{eq:nucleus-u2} and  \ref{eq:nucleus-u2-vector}).

Note that the integrations in equation \ref{eq:nucleus-ui_jth_prism}, are conducted with respect to the variables $(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$, denoting, respectively, the
$x-, \: y-$, and $z-$coordinates of an arbitrary point belonging to the interior
of the $j$th prism (or the $j$th image nucleus).
Finally, the displacement field at the coordinates $x_i$, $y_i$  and $z_i$ due to the pressure change in the whole reservoir can be defined as the sum of the displacements yielded by each prism with constant pressure $\Delta p_{j}$:
\begin{equation}
{\bf{\tilde{u}}}_{i} (x_i,y_i,z_i) = \sum_{j=1}^{M} 
\int_{z_1}^{z_2} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{1}}}_{i} \: \:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j 
\: + \: 
\int_{z_1}^{z_2} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{2}}}_{i} \:\:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j, 
\label{eq:nucleus-ui_M_prisms}
\end{equation}
where $M$ is the number of prisms setting up the reservoir model.

>>>>

The discrete forward modeling to calculate the displacement and stress fields due to a piecewise-constant distribution of the pressure contrast within a reservoir follows the nucleus-of-strain approach. 
We assume that a nucleus of strain represents an infinitesimal reservoir volume 
element. The displacement solution for a single nucleus of strain in a homogeneous elastic semi-infinite medium (equation \ref{eq:displacement-solution}) will be used as an element of the displacement.
Then we calculate the displacement field due to the pressure contrast of a prism by integrating the nucleus of strain over its volume. The same approach is used to 
compute the stress field due to a prism.

@@

By considering the discrete form of equation \ref{eq:displacement-solution}, the displacement vector  $ {\bf{u}}_{ij} \equiv  {\bf{u}}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ at an arbitrary point $(x_{i}, y_{i}, z_{i})$ due to the $j$th nucleus of strain at the coordinates 
$(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ will be calculated by
\begin{equation}
{\bf{u}}_{ij}  = {\bf{u_{1}}}_{ij} + {\bf{u_{2}}}_{ij} \: ,
\label{eq:displacement_ui}
\end{equation}
where ${\bf{u_{1}}}_{ij}$ is the displacement vector at the point $(x_{i}, y_{i}, z_{i})$ due to the $j$th single nucleus in the infinite space and ${\bf{u_{2}}}_{ij}$ 
is the correction of the displacements considering a semi-space (image nucleus solution).
The term ${\bf{u_{1}}}_{ij}$ (equation \ref{eq:displacement_ui}) is given by
\begin{equation}
{\bf{u_{1}}}_{ij} = \frac{A  \: (1 + \nu)}{E} \: {\bf{\nabla}} \bigg( {\frac{1}{{R_1}_{ij}}} \bigg) 
\: \Delta p_{j} \: \:dv_j
\label{eq:nucleus-u1}
\end{equation}
and represents the gradient of the potential
\begin{equation}
\phi_{1} = - \frac{C_m}{4 \pi}  \frac{\Delta p_{j} \: \:dv_j}{ {R_1}_{ij} } \: .
\label{eq:nucleus-phi1}
\end{equation}
The term ${\bf{u_{2}}}_{ij}$ (equation \ref{eq:displacement_ui}) is given by
\begin{equation}
{\bf{u_{2}}}_{ij} = \frac{A  \: (1 + \nu)}{E} \: 
\bigg[ 
(3 - 4\nu){\bf{\nabla}} \bigg(  {\frac{1}{{R_2}_{ij}}} \bigg)     
+ 2 {\bf{\nabla}} \bigg( z \frac{\partial }{\partial z} {\frac{1}{{R_2}_{ij}}} \bigg)  
- 4(1- \nu){\bf{\hat{z}}} {\bf{{\nabla}^{2}}}  \bigg( {\frac{z}{{R_2}_{ij}}} \bigg)
\bigg]   \: \Delta p_{j}  \:dv_j
\label{eq:nucleus-u2}
\end{equation}
and is obtained by applying the operator ${\bf{{\nabla}_{2}}}$ (equation 
\ref{eq:nabla2}) to the image potential
\begin{equation}
\phi_{2} = - \frac{C_m}{4 \pi}  \frac{\Delta p_{j} \: \:dv_j}{ {R_2}_{ij} } \: .
\label{eq:nucleus-phi2}
\end{equation}
In equations \ref{eq:nucleus-u1}--\ref{eq:nucleus-phi2}, 
all derivatives are computed with respect to the coordinates of the point 
$(x_{i}, y_{i}, z_{i})$, 
$\Delta p_j$  is the pressure contrast of the $j$th nucleus, $dv_j$ is an 
infinitesimal element of volume of the $j$th nucleus, $A$ is the constant
\begin{equation}
A = - \frac{C_m E}{4 \pi (1 + \nu)} \: ,
\label{eq:nucleus-A}
\end{equation}
$C_m$ is the uniaxial compaction coefficient (see \citeauthor{Geertsma66}, 
\citeyear{Geertsma66}; \citeauthor{Tempone10}, \citeyear{Tempone10} and 
\citeauthor{Munoz&Roehl17}, \citeyear{Munoz&Roehl17}) 
\begin{equation}
C_m = \frac{1}{E} \: \frac{(1 + \nu) (1  - 2\nu)}{(1-\nu)} \: ,
\label{eq:Cm}
\end{equation}
and $E$ is the Young’s modulus.
In equation \ref{eq:nucleus-phi1}, $ {R_1}_{ij}$ is the distance from the $i$th point $ (x_{i}, y_{i}, z_{i})$ to the $j$th nucleus of strain $(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$, i.e.:
\begin{equation}
{R_1}_{ij} = {\sqrt{(x_{i}- x^{\prime}_{j})^{2} + (y_{i} - y^{\prime}_{j})^{2} + 
(z_{i} - z^{\prime}_{j})^{2}}}.
\label{eq:R1}
\end{equation}
In equation \ref{eq:nucleus-phi2}, $ {R_2}_{ij}$ is the distance from the $i$th point $ (x_{i}, y_{i}, z_{i})$ to the $j$th image nucleus $(x^{\prime}_{j}, y^{\prime}_{j}, - z^{\prime}_{j})$, i.e.:
\begin{equation}
{R_2}_{ij} = {\sqrt{(x_{i}- x^{\prime}_{j})^{2} + (y_{i} - y^{\prime}_{j})^{2} + 
(z_{i} + z^{\prime}_{j})^{2}}} \: .
\label{eq:R2}
\end{equation}
Figure \ref{fig:nucleus_strain} shows a schematic representation of the geometry
of the nucleus of strain problem in a semi-infinite medium. 
The horizontal plane $z = 0$ is called "free surface".
For convenience, the $x-$, $y-$ and $z-$components of the vectors ${\bf{u_{1}}}_{i}$ (equation \ref{eq:nucleus-u1}) and ${\bf{u_{2}}}_{i}$ (equation \ref{eq:nucleus-u2}) 
can be explicitly defined as follows:
\begin{equation}
{\bf{u_{1}}}_{ij} = 
\frac{A  \: (1 + \nu)}{E} \: 
\begin{bmatrix} 
\frac{\partial }{\partial x} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial }{\partial y} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial }{\partial z} {\frac{1}{{R_1}_{ij}}} 
\end{bmatrix}
 \: \Delta p_{j} \: \:dv_j
\label{eq:nucleus-u1-vector}
\end{equation}
and
\begin{equation}
{\bf{u_{2}}}_{ij} = 
\frac{A  \: (1 + \nu)}{E} \: \left\{   (3  - 4 \nu)
\begin{bmatrix} 
  \: \frac{\partial }{\partial x} {\frac{1}{{R_2}_{ij}}}  \\
  \: \frac{\partial }{\partial y} {\frac{1}{{R_2}_{ij}}}  \\
 - \frac{\partial }{\partial z} {\frac{1}{{R_2}_{ij}}}  
\end{bmatrix}
+ 2 \: z_{i}
\begin{bmatrix} 
\frac{\partial^{2}  }{\partial x \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_2}_{ij}}} 
\end{bmatrix}
\right\}
\: \Delta p_{j} \: \:dv_j \: .
\label{eq:nucleus-u2-vector}
\end{equation}

By following \cite{Sharma56} and \cite{Tempone10}, 
the stress vector 
$\mbox{\boldmath$\sigma$}_{ij} \equiv \mbox{\boldmath$\sigma$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ 
at the point $(x_{i}, y_{i}, z_{i})$ due to the $j$th single nucleus of strain buried 
in the half space is given by
\begin{equation}
\mbox{\boldmath$\sigma$}_{ij} = \mbox{\boldmath$\sigma_{1}$}_{ij} + 
\mbox{\boldmath$\sigma_{2}$}_{ij} \: ,
\label{eq:stress_nucleus}
\end{equation}
where 
$\mbox{\boldmath$\sigma_{1}$}_{ij} \equiv \mbox{\boldmath$\sigma_{1}$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ is the stress vector 
at the point $(x_{i}, y_{i}, z_{i})$ due to the $j$th single nucleus in the infinite space and $\mbox{\boldmath$\sigma_{2}$}_{ij} \equiv \mbox{\boldmath$\sigma_{2}$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ is the stress vector 
at the point $(x_{i}, y_{i}, z_{i})$ that gives the correction of the stress due to the 
$j$th image nucleus considering a semi-space. These two vectors are given by
\begin{equation}
\mbox{\boldmath$\sigma_{1}$}_{ij} =
A  \: 
\begin{bmatrix} 
\frac{\partial^{2} }{\partial x \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_1}_{ij}}} 
\end{bmatrix}
\: \Delta p_{j} \: \:dv_j \: ,
\label{eq:nucleus-stress1-vector}
\end{equation}
and 
\begin{equation}
\mbox{\boldmath$\sigma_{2}$}_{ij} = 
A  \: 
\left\{
\begin{bmatrix} 
\frac{\partial^{2}  }{\partial x \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_2}_{ij}}} \\
- \frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_2}_{ij}}}   
\end{bmatrix}
+ 2 \: z_{i}
\begin{bmatrix} 
\frac{\partial^{3}  }{\partial x \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial y \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial z^{3}} {\frac{1}{{R_2}_{ij}}} 
\end{bmatrix}
\right\}
\: \Delta p_{j} \: \:dv_j \: .
\label{eq:nucleus-stress2-vector}
\end{equation}
According to \cite{Sharma56} and \cite{Tempone10}, the Beltrami’s equations \citep{Beltrami} and the equilibrium equations must be satisfied to obtain the contribution of the stress field in the half space. 
Additionally, the following boundary conditions at the free surface ($z_i = 0$) 
must be satisfied
\begin{equation}
\mbox{\boldmath$\sigma$}_{ij} = \mbox{\boldmath$\sigma_{1}$}_{ij} + 
\mbox{\boldmath$\sigma_{2}$}_{ij} = \mbox{ \boldmath$0$},
\label{eq:null_stress}
\end{equation}
where $\mbox{ \boldmath$0$}$ is the null vector.
This condition can be easily verified by adding the vectors 
$\mbox{\boldmath$\sigma_{1}$}_{ij}$ (equation \ref{eq:nucleus-stress1-vector}) and 
$\mbox{\boldmath$\sigma_{2}$}_{ij}$ (equation \ref{eq:nucleus-stress2-vector}) 
computed at any point on the free surface $(x_i, y_i, z_i = 0)$.

@@

We parameterize the reservoir as a grid of juxtaposed right rectangular prisms.
Each grid prism has a constant pressure contrast $\Delta p_{j}$; however, pressure variations among different prisms are allowed. 
To calculate the displacement field produced by the $jth$ prism
at the $i$th coordinates  $(x_i$, $y_i, z_i)$, we integrate the solution deduced for a single nucleus of strain (equation \ref{eq:nucleus-ui_total})
over its volume and obtain
\begin{equation}
{\bf{u}}_{i} (x_i,y_i,z_i) =  
\int_{z_1}^{z_2} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{1}}}_{i} \: \:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j 
\: + \:
\int_{z_1 - 2 \, z_c}^{z_2 - 2 \, z_c} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{2}}}_{i} \:\:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j \: ,
\label{eq:nucleus-ui_jth_prism}
\end{equation}
where the limits of the integrals are defined by the borders and the center 
of the $j$th prism in the following way: $x_1$ and $x_2$ are their south and north borders; $y_1$ and  $y_2$ are their  west and east borders; $z_1$ and $z_2$ are their depths to the top and bottom; and $z_c = \frac{1}{2}(z_1 + z_2)$ is its center depth. 
In equation \ref{eq:nucleus-ui_jth_prism}, the integrations are conducted with respect to the variables $(x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$. 
The integral depending on ${\bf{u_{1}}}_{i}$,
the displacement field due to a single nucleus in the infinite space (equations \ref{eq:nucleus-u1} and  \ref{eq:nucleus-u1-vector}), is conducted over the volume of the $j$th prism. The integral depending on ${\bf{u_{2}}}_{i}$,
the displacement field due to the image nucleus (equations \ref{eq:nucleus-u2} and  \ref{eq:nucleus-u2-vector}), is conducted over the volume of a different prism
symmetrically positioned above the free surface and conveniently called "image prism".
Finally, the displacement field at the coordinates $x_i$, $y_i$  and $z_i$ due to the pressure change in the whole reservoir is defined as the sum of the displacements yielded by each prism with constant pressure $\Delta p_{j}$:
\begin{equation}
{\bf{\tilde{u}}}_{i} (x_i,y_i,z_i) = \sum_{j=1}^{M} 
\int_{z_1}^{z_2} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{1}}}_{i} \: \:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j 
\: + \: 
\int_{z_1}^{z_2} \:\:
\int_{y_1}^{y_2} \:\: 
\int_{x_1}^{x_2} 
{\bf{u_{2}}}_{i} \:\:  dx^{\prime}_j dy^{\prime}_j dz^{\prime}_j, 
\label{eq:nucleus-ui_M_prisms}
\end{equation}
where $M$ is the number of prisms setting up the reservoir model.

%From equation \ref{eq:nucleus-u1-vector},  we can get the   
%$\alpha-$components of the displacement vectors ${\bf{u_{1}}}_{i}$  (the integrand of the first integral of equation \ref{eq:nucleus-ui_M_prisms})
%\begin{equation}
%{u_{1}}_\alpha =  \frac{A  \: (1 + \nu)}{E} \: \Bigg[  \frac{\partial }{\partial \alpha} {\frac{1}{{R_1}_{ij}}} \Bigg] \: \Delta p_{j} \:\:\:  
%\label{eq:u1_alpha_component}
%\end{equation}
%where $\alpha$  belongs to the set of $ x-$ $y-$ and $z-$directions of the Cartesian coordinates system.
